---
title: "Capstone DV"
author: "Elbert H. Nugroho"
date: "9/5/2020"
output: html_document
---

# Covid Case and Death Visualization by Elbert H. Nugroho
## Data Source: Our World in Data
#### URL Link: https://ourworldindata.org/coronavirus-source-data
####Disclaimer: This dashboard was made for R Data Visualization Exercise. The creator of this dashboard was not educated in Epidmiology science, therefore this data visualization maybe flawed and cannot be used for any Epidemiological purposes. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Library Loading 
```{r}
library(gridExtra)
library(patchwork)
library(dplyr)
library(ggplot2)
library(plotly)
library(scales)
library(lubridate)
library(shiny)
library(shinydashboard)
library(glue)
library(flexdashboard)
library(RColorBrewer)
display.brewer.all()
```


#Data Pre-Processing
```{r}
death <- read.csv("DATA/COVID DEATH.csv")
death
```

Preliminary Data Checking
```{r}
glimpse(death)
```

Drop Column:
- biweekly_cases
- biweekly_death

Data Type Transformation:
- date
-location

```{r}
death <- death %>% 
  
  
  #drop column
  select(-biweekly_cases, -biweekly_deaths) %>% 
  #data transformation
  mutate(location = factor(location), date= ymd(date))
death
```

Checking data structure with glimpse

```{r}
glimpse(death)
```

```{r}
table(death$location)
```


Filter data for the first day when a new case was recorded

```{r}
death_new <- death %>% 
  filter(new_cases > 0)
death_new
```

Monthly data aggregation from "death_new"

```{r}
death_new <- death_new %>% 
  mutate( month = month(date, label = T, abbr = F))
death_new
```
```{r}
death_agg <- death_new %>% 
  group_by(location, month) %>% 
  summarise(
    new_cases_monthly = sum(new_cases),
    average_daily_case = mean(new_cases),
    new_deaths_monthly = sum(new_deaths),
    average_daily_deaths = mean(new_deaths),
    total_cases_monthly = max(total_cases),
    total_death_monthly = max(total_deaths),
    mean_weekly_cases = mean(weekly_cases),
    mean_weekly_deaths = mean(weekly_deaths),
    case_fatality_rate_monthly = total_death_monthly / total_cases_monthly *100
  ) %>% 
  ungroup()
  
death_agg
```


Defining NA value as 0

```{r}
death_agg <- death_agg %>% 
  mutate(mean_weekly_cases = ifelse(is.na(mean_weekly_cases),yes = 0, mean_weekly_cases),
         mean_weekly_deaths = ifelse(is.na(mean_weekly_deaths), 0, mean_weekly_deaths)
         
  )

```
Checking for any NA value
```{r}
any(is.na(death_agg))
```
Checking Final data with glimpse()

```{r}
death_agg
```

```{r}
glimpse(death_agg)
```

September Total Case and Death per Country and Filter country for cases above 1000

```{r}
death_agg_sept <- death_agg %>%
          group_by(location) %>%
  summarise(total_cases_by_september = max(total_cases_monthly), total_death_by_september = max(total_death_monthly)) %>% 
  ungroup() %>% 
  filter(total_cases_by_september > 1000)
death_agg_sept
```
Fatality Rate in September

```{r}
death_fatality <- death_agg %>% 
  filter(month == "September" , case_fatality_rate_monthly > 0) %>% 
  select(location, case_fatality_rate_monthly) %>% 
  arrange(case_fatality_rate_monthly)
death_fatality
```


## Data Visualization
Theme Algo

```{r}
theme_algo <- theme(
  panel.background = element_rect(fill = "white"),
  panel.grid.major = element_line(colour = "gray80"),
  panel.grid.minor = element_blank(),
  plot.title = element_text(family = "serif", 
                            size = 18)
)
```
 
Case per Month Trend Line 
```{r}
TL_CPM <- death_agg %>% 
            filter(location %in% c("Indonesia", "Singapore", "Malaysia")) %>% 
            ggplot(aes(month, total_cases_monthly, color = location, group = location,
                       text = glue("Month : {month}
                         Country : {location}
                         Case : {number(total_cases_monthly, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_line() +
            labs(title = "Monthly COVID-19 Case by Country",
                 x = NULL,
                 y = "Number of Cases",
                 color = "Country"
            ) + theme_algo
  
  
ggplotly(TL_CPM, tooltip = "text") 
```
Death per Month Trend line 

```{r}
TL_DPM <- death_agg %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(month, total_death_monthly, color = location, group = location,
                       text = glue("Month : {month}
                         Country : {location}
                         Deaths : {number(total_death_monthly, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_line() +
            labs(title = "Monthly COVID-19 Deaths by Country",
                 x = NULL,
                 y = "Number of Deaths",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_DPM, tooltip = "text")
```

Bar Graph for Cases and Death in September

```{r}
BG_CIS <- death_agg_sept %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(location, total_cases_by_september, fill = location, group = location,
                       text = glue("Country : {location}
                         Cases : {number(total_cases_by_september, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_col() + 
            labs(title = "September COVID-19 Cases by Country",
                 x = NULL,
                 y = "Number of Cases",
                 color = "Country"
            ) + theme_algo

ggplotly(BG_CIS  + scale_fill_brewer(palette = "Reds"), tooltip = "text")
```

```{r}
BG_DIS <- death_agg_sept %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(location, total_death_by_september, color = location, group = location,
                       text = glue("Country : {location}
                         Deaths : {number(total_death_by_september, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_col() +
            labs(title = "September COVID-19 Deaths by Country",
                 x = NULL,
                 y = "Number of Deaths",
                 color = "Country"
            ) + theme_algo

ggplotly(BG_DIS, tooltip = "text")
```
Average Daily Cases and Deaths per Country per Month (May be utilized to see if treatments or government policy for social distancing is effective)

```{r}
TL_ACM <- death_agg %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(month, average_daily_case, color = location, group = location,
                       text = glue("Month : {month}
                         Country : {location}
                         Average Daily Case : {number(average_daily_case, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_line() +
            labs(title = "COVID-19 Daily Case Average by Country",
                 x = NULL,
                 y = "Average Daily Case",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_ACM, tooltip = "text")
```

```{r}
TL_ADM <- death_agg %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(month, average_daily_deaths, color = location, group = location,
                       text = glue("Month : {month}
                         Country : {location}
                         Average Daily Deaths : {number(average_daily_deaths, big.mark = ",", accuracy = 1)}"))) + 
  scale_y_continuous(labels = number_format(big.mark = ","))+ 
            geom_line() +
            labs(title = "COVID-19 Daily Death Average by Country",
                 x = NULL,
                 y = "Average Daily Death",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_ADM, tooltip = "text")
```
Daily Case Trend
```{r}
TL_DC <- death_new %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(date, total_cases, color = location, group = location,
                       text = glue("Date : {date}
                         Country : {location}
                         Cases : {number(total_cases, big.mark = ",", accuracy = 1)}"))) + 
 scale_y_continuous(labels = number_format(big.mark = ",")) +
            scale_x_date(date_breaks = "1 month", labels = date_format(format = "%b"))+
            geom_line() +
            labs(title = "COVID-19 Daily Cases Trend by Country",
                 x = NULL,
                 y = "Number of Daily Cases",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_DC, tooltip = "text")
```
Daily Death Trend

```{r}
TL_DD <- death_new %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(date, total_deaths, color = location, group = location,
                       text = glue("Date : {date}
                         Country : {location}
                         Deaths : {number(total_deaths, big.mark = ",", accuracy = 1)}"))) + 
 scale_y_continuous(labels = number_format(big.mark = ",")) +
            scale_x_date(date_breaks = "1 month", labels = date_format(format = "%b"))+
            geom_line() +
            labs(title = "COVID-19 Daily Death Trend by Country",
                 x = NULL,
                 y = "Number of Daily Deaths",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_DD, tooltip = "text")
```
World Case Rate

```{r}
TL_WC <- death_new %>% 
  filter(location == "World") %>%
  ggplot(aes(date, total_cases, color = location, group = location,
                       text = glue("Date : {date}
                         Number of Cases : {number(total_cases, big.mark = ",", accuracy = 1)}"))) + 
 scale_y_continuous(labels = number_format(big.mark = ",")) +
            scale_x_date(date_breaks = "1 month", labels = date_format(format = "%b"))+
            geom_line() +
            labs(title = "COVID-19 Daily Case Trend",
                 x = NULL,
                 y = "Number of Daily Cases",
                 color = "Country"
            ) + theme_algo

ggplotly(TL_WC, tooltip = "text")
  
```
World Death Rate

```{r}
TL_WD <- death_new %>% 
  filter(location == "World") %>%
  ggplot(aes(date, total_deaths, color = location, group = location,
                       text = glue("Date : {date}
                         Number of Deaths : {number(total_deaths, big.mark = ",", accuracy = 1)}"))) + 
 scale_y_continuous(labels = number_format(big.mark = ",")) +
            scale_x_date(date_breaks = "1 month", labels = date_format(format = "%b"))+
            geom_line() +
            labs(title = "COVID-19 Daily Deaths Trend",
                 x = NULL,
                 y = "Number of Daily Deaths",
                 color = "Country"
            ) + theme_algo
  

ggplotly(TL_WD, tooltip = "text")
```

```{r}
death_fatality
```


```{r}
BG_CFR <- death_fatality %>% 
            filter(location %in% c("Indonesia", "Singapore", "United States")) %>% 
            ggplot(aes(case_fatality_rate_monthly, reorder(location, case_fatality_rate_monthly), fill = location, group = location,
                       text = glue("Country : {location}
                         Case Fatality Rate in Percentage : {case_fatality_rate_monthly}"))) + 
            geom_col() +
            labs(title = "Case Fatality Rate by Country",
                 x = "Case Fatality Rate Percentage",
                 y = "Location",
                 color = "Country"
            ) + theme_algo

ggplotly(BG_CFR + scale_fill_brewer(palette = "OrRd"), tooltip = "text")
```

